---
title: "MA615_MID"
output: pdf_document
date: "2022-11-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environment set up

```{r}
library(tidyverse)
library(magrittr)
library(readxl)
library(dplyr)

strawb <- read_xlsx("/Users/yirong/Desktop/strawberries-2022oct30-a.xlsx", 
                    col_names = T)
```

## Clean Whole data set

```{r}
cnames <- colnames(strawb)

x <- 1:dim(strawb)[2]

T <- NULL

for(i in x){T <- c(T, dim(unique(strawb[i]))[1])}

drop_cols <- cnames[which(T == 1)]

strawb %<>% select(!all_of(drop_cols))

## Arrange the data frame by year and state.
strawb %<>% arrange(Year, State)

strawb %<>% separate(col=`Data Item`,
                     into = c("Strawberries", "type", "items", "units"),
                     sep = ",",
                     fill = "right")

```

## Build 4 subsets
strawb_organic
strawb_non_organic
strawb_chem

```{r}

type_organic <- grep("organic", 
                     strawb$type, 
                     ignore.case = T)

items_organic <- grep("organic", 
                      strawb$items, 
                      ignore.case = T)  ## nothing here

Domain_organic <- grep("organic", 
                       strawb$Domain, 
                       ignore.case = T)


Domain_Category_organic <- grep("organic", 
                                strawb$`Domain Category`, 
                                ignore.case = T)

same <- (intersect(type_organic, Domain_organic)==
           intersect(type_organic, Domain_organic))

length(same)==length(type_organic)

org_rows <- intersect(type_organic, Domain_organic)

strawb_organic <- strawb %>% slice(org_rows, preserve = FALSE)

strawb_non_organic <- strawb %>% filter(!row_number() %in% org_rows)


chem_rows <- grep("BEARING - APPLICATIONS", 
                  strawb_non_organic$type, 
                  ignore.case = T)

chem_rows_1 <- grep("chemical", 
                    strawb_non_organic$Domain, 
                    ignore.case = T)

ins <- intersect(chem_rows, chem_rows_1)

chem_rows_2 <- grep("chemical", 
                    strawb_non_organic$`Domain Category`, 
                    ignore.case = T)

ins_2 <- intersect(chem_rows, chem_rows_2)

strawb_chem <- strawb_non_organic %>% slice(chem_rows, preserve = FALSE)

```
##Clean strawb_organic

```{r}
before_cols = colnames(strawb_organic)
T = NULL
x = length(before_cols)

for(i in 1:x){
  b <- length(unlist(strawb_organic[,i] %>% unique()) )
  T <- c(T,b)
}

drop_cols <- before_cols[which(T == 1)]
strawb_organic %<>% select(!all_of(drop_cols))
after_cols = colnames(strawb_organic)

yy<- grep("MEASURED IN", strawb_organic$items, ignore.case = T)
length(yy)==sum(is.na(strawb_organic$units))

strawb_organic$units<-coalesce(strawb_organic$units, strawb_organic$items)

strawb_organic$units<- str_remove_all(strawb_organic$units, "MEASURED IN ")

strawb_organic$items<- str_remove_all(strawb_organic$items, "- SALES")

strawb_organic$items<- str_remove_all(strawb_organic$items, "MEASURED IN ")

strawb_organic$items[strawb_organic$items==" $"]<-""

strawb_organic$items[strawb_organic$items==" CWT"]<-""

strawb_organic %<>% rename(Markets = items)

strawb_organic %<>% select(Year, State,Markets, units, Value,`CV (%)`)
```
##Above used Haviland's work
##Clean Strawb_non_organic

```{r}
before_cols = colnames(strawb_non_organic)
T = NULL
x = length(before_cols)

for(i in 1:x){
  b <- length(unlist(strawb_non_organic[,i] %>% unique()) )
  T <- c(T,b)
}

drop_cols <- before_cols[which(T == 1)]
strawb_non_organic %<>% select(!all_of(drop_cols))
after_cols = colnames(strawb_non_organic)

strawb_non_organic %<>% separate(col=`Domain Category`, 
                          into = c("dc1", "chem_name"),
                          sep = ":", 
                          fill = "right")
strawb_non_organic$Domain[strawb_non_organic$Domain=="TOTAL"]<-"TOTAL/NOT SPECIFIED"

aa  <- grep("CWT", strawb_non_organic$type,ignore.case = T)
length(aa)
cc<- grep("/", strawb_non_organic$type,ignore.case = T)
length(cc)
bb<-sum(is.na(strawb_non_organic$items))
bb
strawb_non_organic$items<-strawb_non_organic$items %>% 
  replace_na("MEASURED IN $ / CWT")

strawb_non_organic %<>% select(Year, State, items, units, dc1, chem_name, Value)

strawb_non_organic %<>% rename(category = units)

strawb_non_organic$items <- str_remove_all(strawb_non_organic$items, 
                                           "MEASURED IN ")

strawb_non_organic %<>% rename(units = items)

strawb_non_organic$dc1 <- str_remove_all(strawb_non_organic$dc1, "CHEMICAL, ")

strawb_non_organic$dc1 %>% unique()

strawb_non_organic%<>% rename(chem_types = dc1)

strawb_non_organic$chem_name <- str_remove_all(strawb_non_organic$chem_name, "\\(")

strawb_non_organic$chem_name <- str_remove_all(strawb_non_organic$chem_name, "\\)")

strawb_non_organic %<>% separate(col = chem_name,
                          into = c("chem_name","chem_code"),
                          sep = "=",
                          fill = "right"
) 
qq <- grep("ACRE", strawb_non_organic$units, ignore.case = T)
ww<-grep("AVG",strawb_non_organic$category,ignore.case = T)
length(qq)==length(ww)
strawb_non_organic %<>% select(Year, State,units, chem_types, chem_name, chem_code, Value)
dd <- grep("NOT SPECIFIED", strawb_non_organic$chem_types, ignore.case = T)
sum(is.na(strawb_non_organic$chem_name))==length(dd)
strawb_non_organic$chem_name %<>% replace_na("NONE")
strawb_non_organic$chem_code[strawb_non_organic$chem_name == "NONE"]<-"NONE"
strawb_non_organic$chem_code[strawb_non_organic$chem_name== " TOTAL"]<-"TOTAL"
```
##Clean Strawb_chem

```{r}
before_cols = colnames(strawb_chem)
T = NULL
x = length(before_cols)

for(i in 1:x){
  b <- length(unlist(strawb_chem[,i] %>% unique()) )
  T <- c(T,b)
}

drop_cols <- before_cols[which(T == 1)]
strawb_chem %<>% select(!all_of(drop_cols))
after_cols = colnames(strawb_chem)

strawb_chem %<>% separate(col=`Domain Category`, 
                          into = c("dc1", "chem_name"),
                          sep = ":", 
                          fill = "right")


strawb_chem %<>% select(Year, State, items, units, dc1, chem_name, Value)

strawb_chem %<>% rename(category = units)

strawb_chem$items <- str_remove_all(strawb_chem$items, "MEASURED IN ")

strawb_chem %<>% rename(units = items)

strawb_chem$dc1 <- str_remove_all(strawb_chem$dc1, "CHEMICAL, ")

strawb_chem$dc1 %>% unique()

strawb_chem %<>% rename(chem_types = dc1)

strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\(")

strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\)")

strawb_chem %<>% separate(col = chem_name,
                          into = c("chem_name","chem_code"),
                          sep = "=",
                          fill = "right"
) 

strawb_chem %<>% select(Year, State,units, chem_types, chem_name, chem_code, Value)





```
###Build safe_chem and poisons_chem

Poisons Chemicals mentioned in the article:BIFENTHRIN128825,METHYL BROMIDE,CHLOROPICRIN

Safe Chemicals: PHOSPHATE,POTASSIUM BICARBON.
```{r}
poisons_chem<-subset(strawb_chem,chem_name==" BIFENTHRIN "|chem_name==" METHYL BROMIDE "|chem_name==" CHLOROPICRIN ")
safe_chem<-subset(strawb_chem,chem_name==" PHOSPHATE" |chem_name==" POTASSIUM BICARBON. " )
```

##EDA

```{r}
library(ggplot2)
library(hrbrthemes)
##EDA data
strawb_organic_eda<-strawb_organic
strawb_organic_eda$Value<-as.numeric(strawb_organic_eda$Value)
strawb_organic_eda<-na.omit(strawb_organic_eda)

strawb_non_organic_eda<-strawb_non_organic
strawb_non_organic$Value<-as.numeric(strawb_non_organic$Value)
strawb_non_organic_eda<-na.omit(strawb_non_organic_eda)

strawb_chem_eda<-strawb_chem
strawb_chem_eda$Value<-as.numeric(strawb_chem_eda$Value)
strawb_chem_eda<-na.omit(strawb_chem_eda)

poisons_chem_eda<-poisons_chem
poisons_chem_eda$Value<-as.numeric(poisons_chem_eda$Value)
poisons_chem_eda<-na.omit(poisons_chem_eda)

safe_chem_eda<-safe_chem%<>% select(Year, State,units, chem_types, chem_name, Value)
safe_chem_eda$Value<-as.numeric(safe_chem_eda$Value)
safe_chem_eda<-na.omit(safe_chem_eda)
#overall
p1 <- ggplot(data = strawb_organic_eda, mapping = aes(
  x = Year, y = Value, fill = State))
p1 + geom_col()

p2 <- ggplot(data = strawb_non_organic_eda, mapping = aes(
  x = Year, y = Value, fill = State))
p2 + geom_col()

p3 <- ggplot(data = strawb_chem_eda, mapping = aes(
  x = Year, y = Value, fill = State))
p3 + geom_col()

##organic strawb state by state
p4<-ggplot(data = strawb_organic_eda, aes(x = "", y = Value, fill = State)) + geom_bar(stat = "identity") + 
  labs(title = "Organic Strawb Value state by state")+
  coord_polar("y")
p4
p5<-ggplot(data = strawb_non_organic_eda, aes(x = "", y = Value, fill = State)) + geom_bar(stat = "identity") + 
  labs(title = "Non_organic Strawb Value state by state")+
  coord_polar("y")
p5


p7 <- ggplot(poisons_chem_eda, aes(x=Year, y=Value, fill=chem_name)) +
  geom_boxplot()+labs(title = "Poisons Chemical",
                      x = "Year",
                      y = "value")
p7

p8<- ggplot(safe_chem_eda, aes(x=Year, y=Value, fill=chem_name)) +
  geom_boxplot()+labs(title = "Safe Chemical",
                      x = "Year",
                      y = "value")
p8

p9 <- ggplot(poisons_chem_eda, aes(x=Year, y=Value, color=chem_name)) + 
    geom_point(size=6) +
    theme_ipsum()
p9

p10 <- ggplot(safe_chem_eda, aes(x=Year, y=Value, color=chem_name)) + 
    geom_point(size=6) +
    theme_ipsum()
p10

#for cv

p11 <- ggplot(strawb_organic_eda, aes(x=Year, y=Value)) +
  geom_point() +
  geom_smooth(method=lm , color="red", se=FALSE) +
  theme_ipsum()
p11

p12 <- ggplot(strawb_organic_eda, aes(x=Year, y=Value)) +
  geom_point() +
  geom_smooth(method=lm , color="red", se=FALSE) +
  theme_ipsum()
p12

p13 <- ggplot(strawb_non_organic_eda, aes(x=Year, y=Value)) +
  geom_point() +
  geom_smooth(method=lm , color="red", se=FALSE) +
  theme_ipsum()
p13

p14<-ggplot(strawb_organic_eda, aes(fill=Markets, y=Value, x=Year)) + 
    geom_bar(position="dodge", stat="identity")

p14

p15<-ggplot(strawb_non_organic_eda, aes(x=Year, y= chem_types, size = Value)) +
    geom_point(alpha=0.7)
p15

```

